{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-03 thesis/landslide.tex
use cleveref

This chapter will discuss Landslide's outer and inner workings in all their gory detail.
It is intended for the aspiring developer or the ambitious user
and hence unlike other chapters is written in the style of documentation rather than as a report of research results.
The reader interested only in a theoretical introduction to model checking's foundational algorithms,
with detailed and friendly examples to help establish intuitions the later chapters may require,
may skip to \sect{\ref{sec:landslide-algs}}.

@@

This section describes the features of Landslide the average student user should expect to interact with.
% TODO: make more specific section reference
Separate user guides also exist, described in Chapter~\ref{chap:education}.

@@

These scripts accomplish the following setup tasks (among other trivialities):
\begin{itemize}
	\item Copy the user's code into {\tt pebsim/p2-basecode/} or {\tt pebsim/pintos/},
		which contain a pre-annotated Pebbles reference kernel binary or pre-annotated Pintos basecode, respectively.
	\item Build the code in its new location.
	\item Run the instrumentation script on the resulting binary to let Landslide know where all the important functions are
		(see \sect{\ref{sec:landslide-glue}}).
\end{itemize}

@@

The preferred method of invoking Landslide is through Quicksand, the Iterative Deepening wrapper program which has all of Chapter~\ref{chap:quicksand} to itself.
% TODO: add quicksand symlink
This is done via the {\tt ./landslide} script in the top-level directory, which:
\begin{itemize}
	\item Checks if the user needs to run {\tt *-setup.sh} again, in case her source code was more recently updated than the existing annotated build (a common mistake),
	\item Passes its arguments through to {\tt id/landslide-id}, the Quicksand binary,
		and
	\item (If during the student user study,) compresses the resulting log files,
		creates a snapshot tarball of them and the current version of the user's code,
		and sends it to me for nefarious research purposes
		\revisionminor{(see \sect{\ref{sec:education-pebbles-instrumentation}})}.
\end{itemize}

@@

\begin{itemize}
	\item {\tt -C}: enable ``control experiment'' mode, i.e., run only 1 instance of Landslide, with all (non-data-race) preemption points enabled in advance
		% TODO: put a section reference here
	\item {\tt -I}: enable Iterative Context Bounding (requires {\tt -C}, although future work may relax this restriction);
		this generally causes bugs to be found faster should they exist, but degrades completion time
		(\sect{\ref{sec:landslide-icb}})
	\item {\tt -0}: enable Preempt-Everywhere mode (\sect{\ref{sec:quicksand-eval}}, requires {\tt -C})
	\item {\tt -M}: enable Maximal State Space mode,
		which prioritizes the maximal state space to optimize for fast verification,
		abandoning all subset jobs even if they might find bugs faster
		(\sect{\ref{sec:tm-eval}}, incompatible with {\tt -C}).
		According to \sect{\ref{sec:quicksand-soundness}}'s soundness proofs,
		this is equivalent to {\tt -0}
		(and according to my experience, \revisionminor{much} faster as well).
	\item {\tt -H}: use Limited Happens-Before for data-race analysis (\sect{\ref{sec:background-hb}})
		(default for Pebbles kernelspace mode)
	\item {\tt -V}: use vector-clock-based Pure Happens-Before for data-race analysis (\sect{\ref{sec:background-hb}})
		(default for P2/PSU userspace and Pintos modes)
	\item {\tt -X}: support transactional memory (Chapter~\ref{chap:tm})
	\item {\tt -A}: support multiple abort codes during transaction failure (\sect{\ref{sec:tm-implementation}});
		required for testing programs which behave differently under different abort circumstances,
		but impacts the state space size
	\item {\tt -S}: suppress retry aborts during transaction failure (\sect{\ref{sec:tm-implementation}})
	\item {\tt -R}: enable retry-set state space reduction for transactional tests (\sect{\ref{sec:tm-implementation}})
	\item {\tt -P}: support Pintos architecture (enabled automatically when {\tt pintos-setup.sh} is run)
	\item {\tt -4}: support Pebbles architecture (enabled automatically when either {\tt p2-setup.sh} or {\tt psu-setup.sh} is run)
		% TODO: put a section reference
	\item {\tt -e ETAFACTOR}: configure heuristic state space ETA deferring factor (described in detail in {\tt id/option.c})
	\item {\tt -E ETATHRESH}: configure heuristic threshold of state space progress for judging ETA stability (described in detail in {\tt id/option.c})
\end{itemize}

@@

The script {\tt pebsim/landslide} invokes Landslide thus.
It should be run from within the {\tt pebsim/} directory.
When supplied no arguments, it reads configuration options from {\tt pebsim/config.landslide}
(a bash script expected to define certain variables as described in \sect{\ref{sec:landslide-glue}}).
The user may optionally specify a file containing additional config directives
% , such as custom preemption points,
as an argument.\footnote{
Quicksand actually supplies two such files as arguments: one ``static'' config file and one ''dynamic'' config file.
The former contains options which require recompiling Landslide (e.g., whether or not to use ICB is controlled by an {\tt \#ifdef} in Landslide's code),
while the latter contains options which Landslide interprets at runtime (e.g., which preemption points to use).
The static options do not change between Landslide instances in a single Quicksand run,
avoiding long Landslide start-up times.
}
Such supported options are as follows.

@@

\begin{itemize}
	\item {\tt within\_function FUNC} - adds {\tt FUNC} to an allowlist of functions required to appear in the current stack trace before identifying a preemption point (see \sect{\ref{sec:landslide-pps}})
	\item {\tt without\_function FUNC} - as above, but a denylist instead of an allowlist
	\item {\tt within\_user\_function FUNC} - as two above but finds the function in the userspace test program rather than the kernel code.
	\item {\tt without\_user\_function FUNC} - difference to two above same as stated one above.
	\item {\tt data\_race ADDR TID LAST\_CALL CURRENT\_SYSCALL} - specifies a data-race preemption point.
		\begin{itemize}
			\item {\tt ADDR} shall be the code address (in hex) of the racing address,
			{\em before} the execution of which a preemption will be issued.
			\item {\tt TID} indicates a thread ID required to be running for this data race.
				To specify data race PPs across all threads at once, set {\tt FILTER\_DRS\_BY\_TID=0} (see next section).
			\item {\tt LAST\_CALL} indicates a code address required to be the site of the last {\tt call} instruction executed
				(similar to specifying a stack trace, but using a full stack trace here degrades performance too much),
				or 0 to not use this feature.
				From personal experience I found this option rather useless and recommend always supplying 0.
				% TODO: fix this section ref
				For further discussion see \sect{\ref{sec:quicksand-pps}}.
			\item {\tt CURRENT\_SYSCALL} indicates the system call number if a user-space data race comes from within a kernel system call which accesses user memory (Pebbles only).
				Usually 0 (i.e., not in kernel code) but {\tt deschedule}'s system call number is common as well.
		\end{itemize}
	\item {\tt input\_pipe FILENAME} - FIFO file used for receiving messages from Quicksand (e.g. to suspend or resume execution).
		Requires {\tt id\_magic} option to be set (next section below).
		The odds that a human user will find spiritual enlightenment through using this option by hand are infinitesimal.
	\item {\tt output\_pipe FILENAME} - as above but for sending messages.
\end{itemize}

@@

\begin{itemize}
\item {\bf Search algorithm options}
\begin{itemize}
	\item {\tt ICB=0} - enable Iterative Context Bounding (\sect{\ref{sec:landslide-icb}});
		corresponds to {\tt -I} in \sect{\ref{sec:landslide-quicksand-options}}.
	\item {\tt PREEMPT\_EVERYWHERE=0} - enable Preempt-Everywhere mode (\sect{\ref{sec:quicksand-eval}});
		corresponds to {\tt -0} in \sect{\ref{sec:landslide-quicksand-options}}.
	\item {\tt EXPLORE\_BACKWARDS=0} - configure whether, at each newly encountered preemption point,
		to allow the current thread to run first then later upon backtracking to preempt (0),
		or to issue preemptions first and then try continuing the current thread later (1).
		0 tends to produce shorter preemption traces while 1 tends to find bugs faster (\cite[\S{}8.7.1]{landslide}).
		Not compatible with ICB.
\end{itemize}

\item {\bf Memory analysis options}
\begin{itemize}
	\item {\tt PURE\_HAPPENS\_BEFORE=1} - select Pure Happens-Before (1) or Limited Happens-Before (2) (\sect{\ref{sec:background-hb}});
		corresponds to {\tt -V}/{\tt -H} in \sect{\ref{sec:landslide-quicksand-options}}.
	\item {\tt FILTER\_DRS\_BY\_TID=1} - configures whether to use the {\tt TID} parameter of {\tt data\_\allowbreak{}race} described above.
	\item {\tt FILTER\_DRS\_BY\_LAST\_CALL=0} - configures whether to use the {\tt LAST\_CALL} parameter of {\tt data\_race} described above.
	\item {\tt ALLOW\_LOCK\_HANDOFF=0} - configures lockset tracking to permit or disallow a lock taken by one thread to be released by another thread.%
		\footnote{If enabled, accesses performed by the second thread before unlocking will not be considered protected by that lock,
			as Landslide cannot infer what prior event abstractly represented the lock's ownership changing,
			leading to spurious data race reports.
			This could be solved in future work with a new annotation.}
	\item {\tt ALLOW\_REENTRANT\_MALLOC\_FREE=0} - allow two threads to be in {\tt malloc}, {\tt free}, or so on simultaneously without declaring it a bug.%
		\footnote{Used in Pintos, where those functions lock/unlock the heap mutex themselves rather than relying on a wrapper function to do so before invoking them.}
	\item {\tt TESTING\_MUTEXES=0} - configure ``mutex testing'' mode (1),
		in which the data race analysis will not consider a mutex's implementation to be protected by the mutex itself.
		In other words, the mutex's internal memory accesses will be flagged as data races,
		thereby enabling Landslide to verify the mutual exclusion property.
		Normally (0), Landslide assumes mutual exclusion is provided in order to efficiently find data races in the rest of the code.
		Quicksand will automatically set this option for P2s when {\tt -t mutex\_test} is specified.
\end{itemize}

@@

The user may supply custom test cases
for Pebbles (under {\tt pebsim/p2-basecode})
by creating a file in {\tt 410user/progs} and adding it to config.mk as usual,
or for Pintos (under {\tt pebsim/pintos/src/tests/threads})
by creating a file and adding it to both {\tt tests.c} and {\tt Make.tests}.
Tests for the most common interactions during the P2 and Pintos projects are of course already supplied,
as described in \sect{\ref{sec:education-pebbles-tests}} and \sect{\ref{sec:education-pintos-tests}}.

Use of {\tt tell\_landslide()} annotations
\revisionminor{(\sect{\ref{sec:tell-landslide}})}
is not necessary,
although {\tt tell\_landslide\_\allowbreak{}preempt()} and {\tt tell\_landslide\_dump\_stack()}
may optionally be used at the user's convenience.
Additionally, the following ``secret'' annotations
are occasionally used in the pre-supplied test cases
to accomplish several mysterious goals described hereupon.

@@

Many of the supplied P2 test cases invoke the {\tt misbehave} system call
with a mysterious argument (usually {\tt BGND\_BRWN >> FGND\_CYAN})
before the creation of any child threads.
The use of terminal color code constants is of course a red herring of obfuscation,
as the true nature of the Pathos reference kernel's {\tt misbehave} modes
is a closely-guarded secret among 15-410 course staff (\sect{\ref{sec:410-secrecy}}).
The mode in question causes the reference kernel to prioritize scheduling the child thread over the parent
whenever {\tt thread\_fork} is called,
and the target thread over the invoking thread
whenever {\tt make\_runnable} is called,
which are necessary to allow Landslide to recognize a {\tt yield()} preemption point
and be able to run the newly-runnable thread as soon as possible.

To illustrate, consider the following program in Figure~\ref{fig:misbehave-example},
and suppose Landslide is configured to preempt only on mutex API calls
(such as in the first step of Iterative Deepening (\sect{\ref{sec:quicksand-id}})).
Because Landslide ignores all kernel-level synchronization short of context switches when testing user-level code,
if the kernel created the child thread and returned from {\tt thread\_fork}
(the system call underlying {\tt thr\_create()})
without yielding first,
the next preemption point will not occur until {\tt thr\_join()} waits for the child to exit.
Hence, DPOR will erroneously think everything before that {\tt thr\_join()}
happens-before (\sect{\ref{sec:landslide-dpor-hb}}) anything the child does,
and will fail to identify the racing accesses on {\tt x}.

@@

Though Iterative Deepening's soundness (\sect{\ref{sec:quicksand-soundness}})
guarantees all data races will eventually be detected starting from just synchronization preemption points,
it assumes threads becoming runnable counts among those.
In that sense, {\tt misbehave} serves to restore the last synchronization preemptions where they belong.
If at this point the reader wonders why Landslide doesn't just identify
the {\tt thread\_create} and {\tt make\_runnable} system calls in the arbiter itself (\sect{\ref{sec:landslide-arbiter}})
and skip this mysterious user-visible complexity,
they would be right to ask:
I have left it this way for no better reason than to maintain consistency
with the upcoming chapters' experimental environments,
and intend on fixing it in a future update.

@@

When Landslide finds a bug, it produces an execution trace of the particular interleaving of threads that led to the bug.
This takes the form of a two-dimensional table,
with a column for each thread,
and each row representing the continuous execution of one thread between two (not necessarily consecutive) preemption points.
In each row, the cell in the column corresponding to the executed thread will contain a stack trace,
indicating the code location of the preemption point {\em at the end} of that thread transition
(i.e., each stack trace indicates ``this thread ran until it reached the indicated line of code'').
The bug reports are formatted in \revisionminor{HTML}, recommended to be viewed in a web browser.
An example is shown in Figure~\ref{fig:bugreport}.

In addition to the preemption trace, the bug report provides some additional helpful information:
a stack trace of the current thread at the ultimate point when the bug was executed
\revision{(the same as the stack trace in the cell corresponding to that thread in the bottom row of the table)},
a message indicating the nature of the bug encountered,
statistics about the size of the state space,
and optionally additional information about the bug.%
\footnote{For certain types of bugs, not pictured here;
for example, use-after-frees will report separate stack traces
indicating when the corresponding heap block was last allocated and freed.
The intrepid source-diver may find all such cases of extra bug details
by searching for the macro {\tt FOUND\_A\_BUG\_HTML\_INFO} in Landslide's code.}
\revision{\sect{\ref{sec:future-friendly}} discusses how these bug reports might be further improved
by adding more information still,
such as identifying data-race preemption points or listing memory conflicts that occurred during each transition.}

@@

The following annotations are specified in {\tt pebsim/config.landslide} akin to the static configuration options
described in \sect{\ref{sec:landslide-directly}}..
These specify the names of kernel functions, global variables, default values, and so on
which are required to accurately track the kernel's scheduler state:
{\tt CONTEXT\_SWITCH},
{\tt EXEC},
{\tt FIRST\_TID},
{\tt IDLE\_TID},
{\tt INIT\_TID},
{\tt MEMSET},
{\tt PAGE\_FAULT\_WRAPPER},
{\tt READLINE},
{\tt SFREE},
{\tt SHELL\_TID},
{\tt SPURIOUS\_INTERRUPT\_WRAPPER},
\\
{\tt THREAD\_KILLED\_ARG\_VAL},
{\tt THREAD\_KILLED\_FUNC},
{\tt TIMER\_WRAPPER},
{\tt VM\_USER\_COPY},
\\
{\tt VM\_USER\_COPY\_TAIL},
{\tt YIELD}.

@@

		in her test case (\sect{\ref{sec:landslide-friendly-misbehave}}).
	\item {\tt TRUSTED\_THR\_JOIN=0} - if set to 1, forces Landslide to add a happens-before edge (\sect{\ref{sec:landslide-dpor-hb}}) between the exiting of some thread $N$ and the end of any subsequent {\tt thr\_join}$(N)$ call,

@@

The execution tree
%(or state space, or preemption point log, however calling it suits you)
is stored as a chain of preemption point nodes named {\tt struct hax} defined in {\tt tree.h}.
Although the state space of possible interleavings is exponentially-sized,
Landslide does not actually need to store any nodes for execution sequences outside the current variation
(see \sect{\ref{sec:landslide-estimate}} and \sect{\ref{sec:landslide-dpor}} for why),
so the total memory consumption is only $O(n)$ in the number of preemption points in a single program run
(for the test cases used in this thesis, typically 20-1000).
Each {\tt hax} stores the following information:

\begin{itemize}
	\item Basic statistics such as the current instruction pointer, thread ID,
		stack trace of current thread at the moment of preemption,
		depth in the tree, parent node pointer, etc.;
	\item Snapshots of the current state of the scheduler (\sect{\ref{sec:landslide-scheduler}})
		and memory accesses and heaps (\sect{\ref{sec:landslide-memory}});
	\item Simulator-dependent data needed to time travel and resume execution
		from this checkpoint (\sect{\ref{sec:landslide-timetravel}});
	\item List of parent/ancestor nodes with memory conflicts and/or happens-before edges to this one
		for DPOR (\sect{\ref{sec:landslide-dpor}});
	\item Current estimated state space proportion and execution time for the subtree rooted at this node
		(not necessarily fully explored yet) for estimation \sect{\ref{sec:landslide-estimate}};
	\item Whether this point is an {\tt xbegin} invocation
		and if so what {\tt xabort} codes are possible and/or already explored for this transaction
		(\sect{\ref{chap:tm}}).
\end{itemize}

@@

It has one main entrypoint, {\tt sched\_update()}, in which both the state machine is updated and scheduling decisions are made.
The interface also offers helper functions for finding and manipulating {\tt agent}s,
and {\tt sched\_recover()}, which prepares the scheduler to force a new thread to run
after a time travel (\sect{\ref{sec:landslide-timetravel}}).

@@

The second part of {\tt sched\_update()},
conditional on the arbiter identifying preemption points (\sect{\ref{sec:landslide-arbiter}}),
manages timer interrupts to switch to a desired thread.
Whenever a preemption point is reached,
the scheduler first creates a checkpoint in the execution tree (\sect{\ref{sec:landslide-save}}),
asks the arbiter which thread to run next,
and if that thread is different from the current one,
forces the kernel into its timer interrupt handler (\sect{\ref{sec:landslide-interrupce}}).

@@

Landslide invokes the explorer at the end of each execution of the test case,
which analyzes the current branch of the interleaving state space tree
to figure out which alternate branch to try executing next.
Its contents are largely algorithmic rather than architectural
and hence further described in \sect{\ref{sec:landslide-dpor}} and \sect{\ref{sec:landslide-icb}}.

@@

The options specified in \sect{\ref{sec:landslide-directly}}
are handled by a family of \revisionminor{distasteful} shell scripts that live in {\tt pebsim/}.

>>>>

This chapter will discuss Landslide's outer and inner workings in all their gory detail.
It is intended for the aspiring developer or the ambitious user
and hence unlike other chapters is written in the style of documentation rather than as a report of research results.
The reader interested only in a theoretical introduction to model checking's foundational algorithms,
with detailed and friendly examples to help establish intuitions the later chapters may require,
may skip to \cref{sec:landslide-algs}.

@@

This section describes the features of Landslide the average student user should expect to interact with.
% TODO: make more specific section reference
Separate user guides also exist, described in \Cref{chap:education}.

@@

These scripts accomplish the following setup tasks (among other trivialities):
\begin{itemize}
	\item Copy the user's code into {\tt pebsim/p2-basecode/} or {\tt pebsim/pintos/},
		which contain a pre-annotated Pebbles reference kernel binary or pre-annotated Pintos basecode, respectively.
	\item Build the code in its new location.
	\item Run the instrumentation script on the resulting binary to let Landslide know where all the important functions are
		(see \cref{sec:landslide-glue}).
\end{itemize}

@@

The preferred method of invoking Landslide is through Quicksand, the Iterative Deepening wrapper program which has all of \Cref{chap:quicksand} to itself.
% TODO: add quicksand symlink
This is done via the {\tt ./landslide} script in the top-level directory, which:
\begin{itemize}
	\item Checks if the user needs to run {\tt *-setup.sh} again, in case her source code was more recently updated than the existing annotated build (a common mistake),
	\item Passes its arguments through to {\tt id/landslide-id}, the Quicksand binary,
		and
	\item (If during the student user study,) compresses the resulting log files,
		creates a snapshot tarball of them and the current version of the user's code,
		and sends it to me for nefarious research purposes
		\revisionminor{(see \cref{sec:education-pebbles-instrumentation})}.
\end{itemize}

@@

\begin{itemize}
	\item {\tt -C}: enable ``control experiment'' mode, i.e., run only 1 instance of Landslide, with all (non-data-race) preemption points enabled in advance
		% TODO: put a section reference here
	\item {\tt -I}: enable Iterative Context Bounding (requires {\tt -C}, although future work may relax this restriction);
		this generally causes bugs to be found faster should they exist, but degrades completion time
		(\cref{sec:landslide-icb})
	\item {\tt -0}: enable Preempt-Everywhere mode (\cref{sec:quicksand-eval}, requires {\tt -C})
	\item {\tt -M}: enable Maximal State Space mode,
		which prioritizes the maximal state space to optimize for fast verification,
		abandoning all subset jobs even if they might find bugs faster
		(\cref{sec:tm-eval}, incompatible with {\tt -C}).
		According to \cref{sec:quicksand-soundness}'s soundness proofs,
		this is equivalent to {\tt -0}
		(and according to my experience, \revisionminor{much} faster as well).
	\item {\tt -H}: use Limited Happens-Before for data-race analysis (\cref{sec:background-hb})
		(default for Pebbles kernelspace mode)
	\item {\tt -V}: use vector-clock-based Pure Happens-Before for data-race analysis (\cref{sec:background-hb})
		(default for P2/PSU userspace and Pintos modes)
	\item {\tt -X}: support transactional memory (\Cref{chap:tm})
	\item {\tt -A}: support multiple abort codes during transaction failure (\cref{sec:tm-implementation});
		required for testing programs which behave differently under different abort circumstances,
		but impacts the state space size
	\item {\tt -S}: suppress retry aborts during transaction failure (\cref{sec:tm-implementation})
	\item {\tt -R}: enable retry-set state space reduction for transactional tests (\cref{sec:tm-implementation})
	\item {\tt -P}: support Pintos architecture (enabled automatically when {\tt pintos-setup.sh} is run)
	\item {\tt -4}: support Pebbles architecture (enabled automatically when either {\tt p2-setup.sh} or {\tt psu-setup.sh} is run)
		% TODO: put a section reference
	\item {\tt -e ETAFACTOR}: configure heuristic state space ETA deferring factor (described in detail in {\tt id/option.c})
	\item {\tt -E ETATHRESH}: configure heuristic threshold of state space progress for judging ETA stability (described in detail in {\tt id/option.c})
\end{itemize}

@@

The script {\tt pebsim/landslide} invokes Landslide thus.
It should be run from within the {\tt pebsim/} directory.
When supplied no arguments, it reads configuration options from {\tt pebsim/config.landslide}
(a bash script expected to define certain variables as described in \cref{sec:landslide-glue}).
The user may optionally specify a file containing additional config directives
% , such as custom preemption points,
as an argument.\footnote{
Quicksand actually supplies two such files as arguments: one ``static'' config file and one ''dynamic'' config file.
The former contains options which require recompiling Landslide (e.g., whether or not to use ICB is controlled by an {\tt \#ifdef} in Landslide's code),
while the latter contains options which Landslide interprets at runtime (e.g., which preemption points to use).
The static options do not change between Landslide instances in a single Quicksand run,
avoiding long Landslide start-up times.
}
Such supported options are as follows.

@@

\begin{itemize}
	\item {\tt within\_function FUNC} - adds {\tt FUNC} to an allowlist of functions required to appear in the current stack trace before identifying a preemption point (see \cref{sec:landslide-pps})
	\item {\tt without\_function FUNC} - as above, but a denylist instead of an allowlist
	\item {\tt within\_user\_function FUNC} - as two above but finds the function in the userspace test program rather than the kernel code.
	\item {\tt without\_user\_function FUNC} - difference to two above same as stated one above.
	\item {\tt data\_race ADDR TID LAST\_CALL CURRENT\_SYSCALL} - specifies a data-race preemption point.
		\begin{itemize}
			\item {\tt ADDR} shall be the code address (in hex) of the racing address,
			{\em before} the execution of which a preemption will be issued.
			\item {\tt TID} indicates a thread ID required to be running for this data race.
				To specify data race PPs across all threads at once, set {\tt FILTER\_DRS\_BY\_TID=0} (see next section).
			\item {\tt LAST\_CALL} indicates a code address required to be the site of the last {\tt call} instruction executed
				(similar to specifying a stack trace, but using a full stack trace here degrades performance too much),
				or 0 to not use this feature.
				From personal experience I found this option rather useless and recommend always supplying 0.
				% TODO: fix this section ref
				For further discussion see \cref{sec:quicksand-pps}.
			\item {\tt CURRENT\_SYSCALL} indicates the system call number if a user-space data race comes from within a kernel system call which accesses user memory (Pebbles only).
				Usually 0 (i.e., not in kernel code) but {\tt deschedule}'s system call number is common as well.
		\end{itemize}
	\item {\tt input\_pipe FILENAME} - FIFO file used for receiving messages from Quicksand (e.g. to suspend or resume execution).
		Requires {\tt id\_magic} option to be set (next section below).
		The odds that a human user will find spiritual enlightenment through using this option by hand are infinitesimal.
	\item {\tt output\_pipe FILENAME} - as above but for sending messages.
\end{itemize}

@@

\begin{itemize}
\item {\bf Search algorithm options}
\begin{itemize}
	\item {\tt ICB=0} - enable Iterative Context Bounding (\cref{sec:landslide-icb});
		corresponds to {\tt -I} in \cref{sec:landslide-quicksand-options}.
	\item {\tt PREEMPT\_EVERYWHERE=0} - enable Preempt-Everywhere mode (\cref{sec:quicksand-eval});
		corresponds to {\tt -0} in \cref{sec:landslide-quicksand-options}.
	\item {\tt EXPLORE\_BACKWARDS=0} - configure whether, at each newly encountered preemption point,
		to allow the current thread to run first then later upon backtracking to preempt (0),
		or to issue preemptions first and then try continuing the current thread later (1).
		0 tends to produce shorter preemption traces while 1 tends to find bugs faster (\cite[\S{}8.7.1]{landslide}).
		Not compatible with ICB.
\end{itemize}

\item {\bf Memory analysis options}
\begin{itemize}
	\item {\tt PURE\_HAPPENS\_BEFORE=1} - select Pure Happens-Before (1) or Limited Happens-Before (2) (\cref{sec:background-hb});
		corresponds to {\tt -V}/{\tt -H} in \cref{sec:landslide-quicksand-options}.
	\item {\tt FILTER\_DRS\_BY\_TID=1} - configures whether to use the {\tt TID} parameter of {\tt data\_\allowbreak{}race} described above.
	\item {\tt FILTER\_DRS\_BY\_LAST\_CALL=0} - configures whether to use the {\tt LAST\_CALL} parameter of {\tt data\_race} described above.
	\item {\tt ALLOW\_LOCK\_HANDOFF=0} - configures lockset tracking to permit or disallow a lock taken by one thread to be released by another thread.%
		\footnote{If enabled, accesses performed by the second thread before unlocking will not be considered protected by that lock,
			as Landslide cannot infer what prior event abstractly represented the lock's ownership changing,
			leading to spurious data race reports.
			This could be solved in future work with a new annotation.}
	\item {\tt ALLOW\_REENTRANT\_MALLOC\_FREE=0} - allow two threads to be in {\tt malloc}, {\tt free}, or so on simultaneously without declaring it a bug.%
		\footnote{Used in Pintos, where those functions lock/unlock the heap mutex themselves rather than relying on a wrapper function to do so before invoking them.}
	\item {\tt TESTING\_MUTEXES=0} - configure ``mutex testing'' mode (1),
		in which the data race analysis will not consider a mutex's implementation to be protected by the mutex itself.
		In other words, the mutex's internal memory accesses will be flagged as data races,
		thereby enabling Landslide to verify the mutual exclusion property.
		Normally (0), Landslide assumes mutual exclusion is provided in order to efficiently find data races in the rest of the code.
		Quicksand will automatically set this option for P2s when {\tt -t mutex\_test} is specified.
\end{itemize}

@@

The user may supply custom test cases
for Pebbles (under {\tt pebsim/p2-basecode})
by creating a file in {\tt 410user/progs} and adding it to config.mk as usual,
or for Pintos (under {\tt pebsim/pintos/src/tests/threads})
by creating a file and adding it to both {\tt tests.c} and {\tt Make.tests}.
Tests for the most common interactions during the P2 and Pintos projects are of course already supplied,
as described in \cref{sec:education-pebbles-tests} and \cref{sec:education-pintos-tests}.

Use of {\tt tell\_landslide()} annotations
\revisionminor{(\cref{sec:tell-landslide})}
is not necessary,
although {\tt tell\_landslide\_\allowbreak{}preempt()} and {\tt tell\_landslide\_dump\_stack()}
may optionally be used at the user's convenience.
Additionally, the following ``secret'' annotations
are occasionally used in the pre-supplied test cases
to accomplish several mysterious goals described hereupon.

@@

Many of the supplied P2 test cases invoke the {\tt misbehave} system call
with a mysterious argument (usually {\tt BGND\_BRWN >> FGND\_CYAN})
before the creation of any child threads.
The use of terminal color code constants is of course a red herring of obfuscation,
as the true nature of the Pathos reference kernel's {\tt misbehave} modes
is a closely-guarded secret among 15-410 course staff (\cref{sec:410-secrecy}).
The mode in question causes the reference kernel to prioritize scheduling the child thread over the parent
whenever {\tt thread\_fork} is called,
and the target thread over the invoking thread
whenever {\tt make\_runnable} is called,
which are necessary to allow Landslide to recognize a {\tt yield()} preemption point
and be able to run the newly-runnable thread as soon as possible.

To illustrate, consider the following program in \Cref{fig:misbehave-example},
and suppose Landslide is configured to preempt only on mutex API calls
(such as in the first step of Iterative Deepening (\cref{sec:quicksand-id})).
Because Landslide ignores all kernel-level synchronization short of context switches when testing user-level code,
if the kernel created the child thread and returned from {\tt thread\_fork}
(the system call underlying {\tt thr\_create()})
without yielding first,
the next preemption point will not occur until {\tt thr\_join()} waits for the child to exit.
Hence, DPOR will erroneously think everything before that {\tt thr\_join()}
happens-before (\cref{sec:landslide-dpor-hb}) anything the child does,
and will fail to identify the racing accesses on {\tt x}.

@@

Though Iterative Deepening's soundness (\cref{sec:quicksand-soundness})
guarantees all data races will eventually be detected starting from just synchronization preemption points,
it assumes threads becoming runnable counts among those.
In that sense, {\tt misbehave} serves to restore the last synchronization preemptions where they belong.
If at this point the reader wonders why Landslide doesn't just identify
the {\tt thread\_create} and {\tt make\_runnable} system calls in the arbiter itself (\cref{sec:landslide-arbiter})
and skip this mysterious user-visible complexity,
they would be right to ask:
I have left it this way for no better reason than to maintain consistency
with the upcoming chapters' experimental environments,
and intend on fixing it in a future update.

@@

When Landslide finds a bug, it produces an execution trace of the particular interleaving of threads that led to the bug.
This takes the form of a two-dimensional table,
with a column for each thread,
and each row representing the continuous execution of one thread between two (not necessarily consecutive) preemption points.
In each row, the cell in the column corresponding to the executed thread will contain a stack trace,
indicating the code location of the preemption point {\em at the end} of that thread transition
(i.e., each stack trace indicates ``this thread ran until it reached the indicated line of code'').
The bug reports are formatted in \revisionminor{HTML}, recommended to be viewed in a web browser.
An example is shown in \Cref{fig:bugreport}.

In addition to the preemption trace, the bug report provides some additional helpful information:
a stack trace of the current thread at the ultimate point when the bug was executed
\revision{(the same as the stack trace in the cell corresponding to that thread in the bottom row of the table)},
a message indicating the nature of the bug encountered,
statistics about the size of the state space,
and optionally additional information about the bug.%
\footnote{For certain types of bugs, not pictured here;
for example, use-after-frees will report separate stack traces
indicating when the corresponding heap block was last allocated and freed.
The intrepid source-diver may find all such cases of extra bug details
by searching for the macro {\tt FOUND\_A\_BUG\_HTML\_INFO} in Landslide's code.}
\revision{\cref{sec:future-friendly} discusses how these bug reports might be further improved
by adding more information still,
such as identifying data-race preemption points or listing memory conflicts that occurred during each transition.}

@@

The following annotations are specified in {\tt pebsim/config.landslide} akin to the static configuration options
described in \cref{sec:landslide-directly}..
These specify the names of kernel functions, global variables, default values, and so on
which are required to accurately track the kernel's scheduler state:
{\tt CONTEXT\_SWITCH},
{\tt EXEC},
{\tt FIRST\_TID},
{\tt IDLE\_TID},
{\tt INIT\_TID},
{\tt MEMSET},
{\tt PAGE\_FAULT\_WRAPPER},
{\tt READLINE},
{\tt SFREE},
{\tt SHELL\_TID},
{\tt SPURIOUS\_INTERRUPT\_WRAPPER},
\\
{\tt THREAD\_KILLED\_ARG\_VAL},
{\tt THREAD\_KILLED\_FUNC},
{\tt TIMER\_WRAPPER},
{\tt VM\_USER\_COPY},
\\
{\tt VM\_USER\_COPY\_TAIL},
{\tt YIELD}.

@@

		in her test case (\cref{sec:landslide-friendly-misbehave}).
	\item {\tt TRUSTED\_THR\_JOIN=0} - if set to 1, forces Landslide to add a happens-before edge (\cref{sec:landslide-dpor-hb}) between the exiting of some thread $N$ and the end of any subsequent {\tt thr\_join}$(N)$ call,

@@

The execution tree
%(or state space, or preemption point log, however calling it suits you)
is stored as a chain of preemption point nodes named {\tt struct hax} defined in {\tt tree.h}.
Although the state space of possible interleavings is exponentially-sized,
Landslide does not actually need to store any nodes for execution sequences outside the current variation
(see \cref{sec:landslide-estimate} and \cref{sec:landslide-dpor} for why),
so the total memory consumption is only $O(n)$ in the number of preemption points in a single program run
(for the test cases used in this thesis, typically 20-1000).
Each {\tt hax} stores the following information:

\begin{itemize}
	\item Basic statistics such as the current instruction pointer, thread ID,
		stack trace of current thread at the moment of preemption,
		depth in the tree, parent node pointer, etc.;
	\item Snapshots of the current state of the scheduler (\cref{sec:landslide-scheduler})
		and memory accesses and heaps (\cref{sec:landslide-memory});
	\item Simulator-dependent data needed to time travel and resume execution
		from this checkpoint (\cref{sec:landslide-timetravel});
	\item List of parent/ancestor nodes with memory conflicts and/or happens-before edges to this one
		for DPOR (\cref{sec:landslide-dpor});
	\item Current estimated state space proportion and execution time for the subtree rooted at this node
		(not necessarily fully explored yet) for estimation \cref{sec:landslide-estimate};
	\item Whether this point is an {\tt xbegin} invocation
		and if so what {\tt xabort} codes are possible and/or already explored for this transaction
		(\cref{chap:tm}).
\end{itemize}

@@

It has one main entrypoint, {\tt sched\_update()}, in which both the state machine is updated and scheduling decisions are made.
The interface also offers helper functions for finding and manipulating {\tt agent}s,
and {\tt sched\_recover()}, which prepares the scheduler to force a new thread to run
after a time travel (\cref{sec:landslide-timetravel}).

@@

The second part of {\tt sched\_update()},
conditional on the arbiter identifying preemption points (\cref{sec:landslide-arbiter}),
manages timer interrupts to switch to a desired thread.
Whenever a preemption point is reached,
the scheduler first creates a checkpoint in the execution tree (\cref{sec:landslide-save}),
asks the arbiter which thread to run next,
and if that thread is different from the current one,
forces the kernel into its timer interrupt handler (\cref{sec:landslide-interrupce}).

@@

Landslide invokes the explorer at the end of each execution of the test case,
which analyzes the current branch of the interleaving state space tree
to figure out which alternate branch to try executing next.
Its contents are largely algorithmic rather than architectural
and hence further described in \cref{sec:landslide-dpor} and \cref{sec:landslide-icb}.

@@

The options specified in \cref{sec:landslide-directly}
are handled by a family of \revisionminor{distasteful} shell scripts that live in {\tt pebsim/}.
